### 2023-10-26 README.qmd
final version of readme

---
title: "Working memory performance in adverse environments: Enhanced, impaired, or intact?"
output: github_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=F, echo=F}
library(tidyverse)
devtools::load_all("D:/repositories/projectlog")
#source('scripts/dependencies.R')

knitr::opts_chunk$set("ft.shadow" = FALSE, echo = F, message = F, warning = F)

```
<br>
This repository contains data, code, and output for a project entitled "Working memory performance in adverse environments: Enhanced, impaired, or intact?". A Stage 1 snapshot of the project has been submitted to [Peer Community in Registered Reports (PCI-RR)](https://rr.peercommunityin.org/).

# Project History

Below is an overview of all the major milestones of this project, including a description and links.
The links take you to the historic version of the project that corresponds to the milestone.

```{r}
proj_history <-
  gert::git_tag_list() |>
  dplyr::rename(
    milestone = name,
    milestone_hash = commit) |>
  mutate(
    commit_hash = (\(x) purrr::pmap_chr(list(milestone, ref, milestone_hash), .f = function(milestone, ref, milestone_hash) {
      gert::git_commit_info(milestone_hash)$id
    }))()
  ) |>
  left_join(
    gert::git_log() |>
      rename(commit_hash = commit)
  ) |>
  mutate(
    milestone_link = paste0(get_git_url(), '/releases/tag/', milestone),
    link = commit_hash,
    commit_link = paste0(get_git_url(), '/tree/', commit_hash)
  ) |> 
  separate(message, into = c("message", "code"), sep = "code") |> 
  mutate(
    code = str_replace_all(code, "\\|\\>", "\\|\\>\\\n"),
    message = str_remove(message, "\\\nobject_hash.*$"),
    milestone = str_c("[", milestone, "]", "(www.google.com)")
  )

knitr::knit_print(flextable::flextable(proj_history) |> 
  flextable::autofit() |> 
  flextable::compose(
    j = "milestone",
    value = flextable::as_paragraph(
      flextable::hyperlink_text(x = milestone, url = milestone_link)
    )
  ) |> 
  flextable::compose(
    j = "link",
    value = flextable::as_paragraph(
      flextable::hyperlink_text(x = "Click here", url = commit_link)
    )
  ) |> 
  flextable::delete_columns(c("ref", "milestone_hash", "commit_hash", "files", "merge", "milestone_link", "commit_link")))
```

### Contact

For questions or comments, feel free to contact me at p.c.s.vermeent@gmail.com.