### 2024-01-22 README.Rmd
remove milestones for now

---
title: ""
output: 
  html_document:
    template: assets/template.html
    css: assets/style.css
    keep_md: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=F, echo=F}
library(tidyverse)
library(glue)
devtools::load_all("D:/repositories/projectlog")
#source('scripts/dependencies.R')

knitr::opts_chunk$set("ft.shadow" = FALSE, echo = F, message = F, warning = F)

```
<br>
This repository contains data, code, and output for a project entitled "Working memory performance in adverse environments: Enhanced, impaired, or intact?". A Stage 1 snapshot of the project has been submitted to [Peer Community in Registered Reports (PCI-RR)](https://rr.peercommunityin.org/).

Do you want to download or clone the materials for this project? Go to [https://github.com/stefanvermeent/liss_wm_profiles_2023](https://github.com/stefanvermeent/liss_wm_profiles_2023).

## Directory Structure {#structure}

The names of each folder are intended to be self-explanatory.
There are eight top-level folders to organize the inputs and outputs of this project:

1.  [`Manuscript`](https://stefanvermeent.github.io/abcd_ddm/registered_report/README.html): The Registered Report written in Quarto.
2.  [`Supplement`](https://stefanvermeent.github.io/abcd_ddm/supplement/README.html): a supplemental text (to be submitted with the manuscript) documenting all secondary analyses in detail.
3.  [`Scripts`](https://stefanvermeent.github.io/abcd_ddm/scripts/README.html): R-scripts that read, analyze, and produce all analysis objects.
4.  [`Data`](https://stefanvermeent.github.io/abcd_ddm/data/README.html): Folder in which real LISS data can be placed to make the analyses fully reproducible. Note that we cannot openly share the raw data on the open repository. The folder contains synthetic data to facilitate computational reproducibility in the absence of LISS access.
5.  [`Analysis Objects`](https://stefanvermeent.github.io/abcd_ddm/analysis_objects/README.html): Folder containing all analysis objects.
6.  [`Codebooks`](https://stefanvermeent.github.io/abcd_ddm/codebooks/README.html): lists of variable names, labels, and value labels (where applicable).

Click on each of the folders to get more details.

## Overview of project milestones

Below is an overview of all the project milestones, such as first-time data access, submissions, and revisions.
Data access events were automatically captured using custom code, which over the course of this project was collected in the R package `projectlog` [https://stefanvermeent.github.io/projectlog/](https://stefanvermeent.github.io/projectlog/).
For more information about how tracking worked, Go to the [Open science workflow tab](https://stefanvermeent.github.io/abcd_ddm/opensci_workflow/README.html).