### 2024-01-22 analysis_objects/README.Rmd
READMEs for the various folders

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Analysis objects

This folder contains all analysis objects, i.e., the outputs of the various data processing scripts.
Similar to the data, analysis objects containing individual-level data cannot be shared openly on GitHub.
These analysis objects can be created either by using the simulated data files or by requesting access to the LISS data (see [how to reproduce this repository](#reproduce)).