### 2024-01-22 codebooks/README.Rmd

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Codebooks

Below you can find the codebooks of the data sets used for the primary analyses. To download them, go [here](https://github.com/stefanvermeent/liss_wm_profiles_2023/tree/main/codebooks).

## ddm_data

```{r, include = FALSE, echo = FALSE, warning = FALSE, message=TRUE}
library(readr)
ddm_data <- read_csv("ddm_data.csv")
```

```{r}
ddm_data
```

## iv_data

```{r, include = FALSE, echo = FALSE, warning = FALSE, message=TRUE}
iv_data <- read_csv("iv_data.csv")
```

```{r}
iv_data
```

## test_set

```{r, include = FALSE, echo = FALSE, warning = FALSE, message=TRUE}
test_set <- read_csv("test_set.csv")
```

```{r}
test_set
```