### 2024-01-23 materials/README.Rmd
README file for materials

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

## Materials {#materials}

This folder contains all the materials of the cognitive tasks.
All tasks were programmed in [JsPsych 7.3.2](https://www.jspsych.org/7.3/).

1.  [`0_pilot/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/0_pilot): Folder containing the pilot materials for the Operation Span Task and the Binding-Updating Task.
2.  [`0_pilot2/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/0_pilot2): Folder containing the pilot materials for the Rotation Span Task.
3.  [`1_liss/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/1_liss): Folder containing the final materials of all three tasks as used in the main study.

In each of these study folders, you will find several files and folders:

1.  `css/`. Handles the visual appearance of the cognitive tasks.
2.  `functions/`. Several helper functions to create stimuli and/or timelines.
3.  `plugins/`. Custom JsPsych plugins to handle trial behavior that is not included in the default JsPsych library.
4.  `global_variables.js`. Global functions for control across studies, e.g., full-screen mode, turning off the cursor.
5.  `index.html`. Script which combines the code of all three tasks, handles counterbalancing, and initiates the tasks.

Within each study folder, each task also has it's own folder (`ospan`, `rspan`, or `updating_binding`.)
In each folder, you will find two files:
`0_[task]_instructions.js` contains the Javascript code for all task instructions.
`1_[task]_procedure.js` contains Javascript code defining the stimuli and trials for both practice and test blocks.

**Run tasks locally.** To run the tasks locally on your device, navigate to one of the three main folders as listed above, and open the `index.html` file in a browser. 
This will automatically start the cognitive tasks (in counterbalanced order).

**Run tasks online.** The tasks are hosted online at [https://liss-wm-profiles-2023-main.netlify.app/](https://liss-wm-profiles-2023-main.netlify.app/).
Navigating to this url will automatically start the tasks.