### 2024-01-23 data/README.Rmd
commit simulated data

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

## Data {#data}

This folder contains all the data that serves as input for the analysis scripts.
LISS data cannot be shared on open repositories.

**Full reproduction**. In order to fully reproduce the analysis described in the manuscript, the interested reader needs to have their own personal access to the LISS data archive.
After obtaining access, the required data files can be placed in the `data` folder (see [how to reproduce this repository](https://stefanvermeent.github.io/liss_wm_profiles_2023/) for more information).

**Computational reproduction**. To facilitate computational reproducibility, we provide simulated data files.
These files contain simulated data with the same variables as the original data.
They are generated under `scrips/1_sim_pipeline` and can be used as input to the data analysis functions.
Note that, because the values are not identical to the real data, the output will deviate from the statistics described in the manuscript.